\section{ADR-0002: Engineering Shadow Scope}

\subsection*{Status}

Accepted

\subsection*{Context}

Initial discussions focused primarily on supporting RTL implementation and debugging.

Further analysis of the RTL engineering workflow revealed that the majority of engineering decisions are made before implementation begins.

Important activities such as theory building, literature review, architecture exploration, requirement analysis, and design planning produce critical engineering knowledge that is not captured by existing tools.

Restricting the product to implementation support would ignore a substantial portion of the engineering lifecycle.

\subsection*{Decision}

Engineering Shadow shall support the complete RTL engineering lifecycle.

The system shall begin assisting from project assignment and continue throughout learning, research, architecture design, implementation, verification, optimization, integration, and project presentation.

\subsection*{Rationale}

Engineering context is created continuously throughout the project lifecycle rather than only during coding.

Supporting the complete lifecycle enables the system to preserve engineering reasoning from the earliest stages of the project and maintain consistent project understanding as the design evolves.

\subsection*{Alternatives Considered}

\begin{enumerate}
    \item RTL coding assistant only.
    \item Verification assistant only.
    \item Debugging assistant only.
\end{enumerate}

These alternatives were rejected because they fail to preserve the reasoning that exists before implementation and therefore cannot maintain complete project understanding.

\subsection*{Consequences}

Future product capabilities shall be evaluated across the entire RTL engineering lifecycle rather than only during implementation.

The Engineering Shadow shall continuously update its understanding as the project evolves through each engineering stage.