\section{ADR-0003: Project Understanding as the Core Product Capability}

\subsection*{Status}

Accepted

\subsection*{Context}

General-purpose AI assistants provide responses based primarily on the information available within the current interaction.

As engineering projects evolve, important context including project objectives, architectural evolution, engineering decisions, design constraints, previous experiments, verification results, and guide feedback becomes fragmented across multiple tools and conversations.

Engineers are therefore required to repeatedly reconstruct project context before meaningful assistance can be provided.

\subsection*{Decision}

Engineering Shadow shall maintain a continuously evolving understanding of the engineering project.

All assistant capabilities, including documentation generation, design explanation, research assistance, debugging support, implementation guidance, and engineering recommendations, shall operate on this shared project understanding.

\subsection*{Rationale}

The primary value of Engineering Shadow is context-aware engineering assistance rather than isolated question answering.

Maintaining project understanding enables the system to provide recommendations that remain consistent with project objectives, design constraints, previous decisions, architectural evolution, and historical engineering reasoning.

This shared understanding forms the foundation upon which all higher-level product capabilities are built.

\subsection*{Alternatives Considered}

\begin{enumerate}
    \item Session-based conversational memory.
    \item Prompt-based context supplied by the engineer.
    \item Static documentation as the primary knowledge source.
\end{enumerate}

These alternatives require repeated manual effort and do not maintain a continuously evolving representation of project understanding.

\subsection*{Consequences}

Project Understanding becomes the central capability of Engineering Shadow.

Future components including Engineering Memory, Decision Memory, Project State, reasoning modules, documentation generation, and engineering guidance shall operate on this shared understanding rather than isolated user interactions.